Synthesizing Verified Components for Cyber Assured Systems Engineering \thanks{DISTRIBUTION STATEMENT A. Approved for public release.}

### 2022-03-21 paper/high-assurance-contracts.tex
added definition of evaluation, plus related discussion

Generally, AGREE specifications \emph{do not} describe the computation
that a component performs. This is entirely by design: AGREE is
intended to reason about component behavior solely at the
specification level. However, the syntax of AGREE specifications
provides enough expressiveness to support the notion of an
AGREE \emph{code contract}: a component contract from which SPLAT can
extract an implementation.

@@

\begin{definition}[Code contract]
A component of the form $(I_c,O_c,A_c,P_c)$ is a \emph{code contract}
if $E \cup G \subset P_c$ and $E$ is a non-empty list of \konst{eq}
statements and $G$ is a set of code guarantees, one for each element
of $O_c$
\end{definition}

The AGREE code contracts for the filter and monitor are shown in
\figref{fig:filter} and \figref{fig:monitor}. In general, the
\konst{eq}-statements of a code contract keep track of
state and provide data values from which to compute values to place on
the outputs.

\subsubsection{Monitor specification}
 For the monitor, BriefCASE generates everything but the
\konst{eq}-statements in \linesref{line:mon-assume}{line:mon-policy}.  As with
the filter, the guarantees completely define the output, as required
for synthesis, in terms of the policy and the alert.  The
\texttt{alert} variable defined on \lineref{line:mon-alert} depends
not just on the policy but also on configuration details supplied when
the component is specified.  At configuration
time, \texttt{is\_latched} has been set to \konst{true}, thus making
the alert persistent, meaning that once the alert is raised, it is
always raised; otherwise, it is the complement of the policy value in
the current step.

>>>>

Generally, AGREE specifications do not describe the computation
that a component performs. This is entirely by design: AGREE is
intended to reason about component behavior solely at the
specification level. However, the syntax of AGREE specifications
provides enough expressiveness to support the notion of a
\emph{code contract}: a component contract from which an implementation
can be extracted.

@@

\begin{definition}[Code contract]
A component of the form $(I_c,O_c,A_c,P_c)$ is a \emph{code contract}
if $\mathit{Eqs} \cup G \subseteq P_c$, where
\[\mathit{Eqs} = \set{v_1 = e_1, \cdots , v_n = e_n} \] is a non-empty set
of \konst{eq} statements and $G$ is the set of output guarantees, one for
each element of $O_c$. In the interpretation as code, the order of
elements of $\mathit{Eqs}$ is important, and is simply taken to be the
occurrence order of the \konst{eq} statements in the syntax. Thus we
will work with the
\emph{list} of equations $\mathit{Eqs} = [v_1 = e_1; \cdots ; v_n = e_n]$.
\end{definition}

Code contracts for the filter and monitor are shown in
\figref{fig:filter} and \figref{fig:monitor}.\footnote{KLS: do we discuss wellformedness anywhere?} In general, the
\konst{eq}-statements of a code contract keep track of
state and provide data values from which to compute values to place on
the outputs.

\subsubsection{Monitor specification}
BriefCASE generates everything but the
\konst{eq}-statements in \linesref{line:mon-assume}{line:mon-policy} of
\figref{fig:monitor}.  As with the filter, the guarantees
completely define the output, as required for synthesis, in terms of
the policy and the alert.  The
\texttt{alert} variable defined on \lineref{line:mon-alert} depends
not just on the policy but also on configuration details supplied when
the component is specified.  At configuration
time, \texttt{is\_latched} has been set to \konst{true}, thus making
the alert persistent, meaning that once the alert is raised, it is
always raised; otherwise, it is the complement of the policy value in
the current step.

### 2022-03-21 paper/intro.tex
added definition of evaluation, plus related discussion

CASE was scoped to address common cyber-vulnerabilities, such as
overflow, lack of input validation, supply-chain issues, and identity
verification; however, other cyber-vulnerabilities such as
side-channel attacks and denial of service are yet dealt with in our
work.  Research in CASE has not invented a new type of filter or
monitor, in terms of capability; instead, our contribution is in the
automated synthesis of these components from a formal specification,
and a means to show that the synthesis is correct.

>>>>

Our approach currently applies to common cyber-vulnerabilities, such
as overflow, lack of input validation, supply-chain issues, and
identity verification; however, other cyber-vulnerabilities such as
side-channel attacks and denial of service are not yet dealt with in
our work.  Here, we do not report on the invention of a new type of
filter or monitor, in terms of capability; instead, our contribution
is in the automated synthesis of security-improving components from
formal specifications, and a means to show that the synthesis is
correct.

### 2022-03-21 paper/semantics.tex
added definition of evaluation, plus related discussion

The syntax of AGREE is essentially that of quantifier-free first order
logic (FOL) supplemented with a few temporal operators. The terms
(\emph{e}) are the usual arithmetic expressions built from variables
and numeric literals, while formulas (\emph{b}) are built using the
standard logical connectives from atomic formulas (\emph{a}) based on
the usual comparison operators.
\[
\begin{array}{rcl}
e & ::= & v \mid c \mid e \;\set{+,*,/}\; e \\
a & ::= & e\; \set{=,<,\leq}\; e \\
b & ::= & v \mid c \mid a \mid \neg b
            \mid b \; \set{\land,\lor,\imp,\iff}\; b
\end{array}
\]

There is a conditional, $\itelse{b}{(-)}{(-)}$, for both terms and
formulas. There are also temporal operators $\konst{pre}(-)$, delay
$(-) \to (-)$, and $\konst{Historically}(-)$.

The semantics of terms and formulas is in terms of \emph{streams of
values}. Values encompass at least booleans and numbers, but can be
readily extended to include records and arrays. A value stream is a
total function from natural numbers to values:
\[
 \konst{stream} = \mathbb{N} \to \konst{value}
\]
Given an \emph{environment} $E : \konst{name} \mapsto \konst{stream}$
binding variable names to value streams, the semantics $\sem{-}$ of terms and
formulas, omitting temporal operators, is standard.

\[
\begin{array}{rcl}
\sem{v}_t & = & E(v)(t) \\
\sem{c}_t & = & c \\
\sem{e_1 + e_2}_t & = & \sem{e_1}_t + \sem{e_2}_t \\
   & \cdots & \\
\sem{b_1 \land b_2}_t & = & \sem{b_1}_t \land \sem{b_2}_t \\
   & \cdots & \\
\end{array}
\]

>>>>

The syntax of AGREE is essentially that of quantifier-free first order
logic (FOL) supplemented with a few temporal operators. The terms
(\emph{e}) are the usual arithmetic expressions built from variables
and numeric literals, while formulas (\emph{b}) are built using the
standard logical connectives from atomic formulas (\emph{a}) based on
the usual comparison operators.
\[
\begin{array}{rcl}
e & ::= & v \mid c \mid e \;\set{+,*,/}\; e \\
a & ::= & e\; \set{=,<}\; e \\
b & ::= & v \mid c \mid a \mid \neg b
            \mid b \; \set{\land,\lor,\imp,\iff}\; b
\end{array}
\]

There is a conditional, $\itelse{b}{(-)}{(-)}$, for both terms and
formulas. There are also temporal operators $\konst{pre}(-)$, delay
$(-) \to (-)$, and $\konst{Hist}(-)$.

The semantics of terms and formulas is in terms of \emph{streams of
values}. Values encompass at least booleans and numbers, but can be
readily extended to include records and arrays. A value stream is a
total function from time (natural numbers) to values:
\[
 \konst{stream} = \mathbb{N} \to \konst{value}
\]
Given an \emph{environment} $E : \konst{name} \mapsto \konst{stream}$
binding variable names to value streams, the semantics $\sem{-}^E_t$
of terms and formulas, omitting the temporal operators, is
conventional. The value of a variable $v$ at time $t$ is found by
looking up the stream bound to $v$ in $E$ (call it $s$) and extracting
the value of $s$ at time $t$.

\[
\begin{array}{rcl}
\sem{v}^E_t & = & E(v)(t) \\
\sem{c}^E_t & = & c \\
\sem{e_1 + e_2}^E_t & = & \sem{e_1}^E_t + \sem{e_2}^E_t \\
   & \cdots & \\
\sem{b_1 \land b_2}^E_t & = & \sem{b_1}^E_t \land \sem{b_2}^E_t \\
   & \cdots & \\
\end{array}
\]

The temporal operators deal with time in more significant ways. For
example, they are used to define higher-level operators from PTLTL,
such as \konst{Once} and \konst{Since}. In code contracts, the
temporal operators are essential for manipulating state variables
between component invocations.

\[
\begin{array}{rcl}
\sem{\konst{pre}(e)}^E_t & = & \sem{e}^E_{t-1}, \mathrm{when}\ t > 0 \\
\sem{e_1 \to e_2}^E_t & = & \itelse{t=0}{\sem{e_1}^E_0}{\sem{e_2}^E_t} \\
\sem{\konst{Hist}(b)}^E_t & = & \forall n \leq t.\; \sem{b}^E_n
\end{array}
\]

Code generation is a multi-step transformation that starts with a code
contract and ends with an executable. The first step maps the contract
to a code-focused representation. Assume given code contract
$(I,O,A,\mathit{Eqs} \cup G \cup P)$, environment $E$, and time
$t$. The evaluation of $\mathit{Eqs}$ followed by the evaluation of
$G$ results in a new environment $E'$ where the stream values for
state variables and outputs have been computed for time $t$. The step
from $E$ to $E'$ is one full cycle in the repeated evaluation of the
component.

\begin{definition}[Evaluation]
We overload existing notation and write the evaluation of
$\mathit{Eqs}$ and then $G$ in environment $E$ at time $t$ as
$\sem{\mathit{Eqs}\cdot G}^E_t$. Each $v = e \in {\Eqs}$ is treated as
an environment transformation
\[
 \sem{v = e}^E_t = E[v(t) \mapsto\sem{e}^E_t] \ .
\]
modifying $E$ so that stream $v$ has value $\sem{e}^E_t$ at time $t$.
{\Eqs} is evaluated by folding the transformation left-to-right
through it. The transformation is also applied to the list of output
guarantees $G$, computing the values on output ports. The details are
omitted, being a bit messy because of the different kinds of output
ports available.
\end{definition}

\begin{definition}[Code contract correctness]
Assume given code contract $(I,O,A,\mathit{Eqs} \cup G \cup P)$ and
environment $E$.  The contract is \emph{correct}, if for all $t$,
whenever the assumptions (historically) hold in $E$ and evaluation
steps from $E$ to $E'$, then the guarantees hold in $E'$:
\[
\sem{\konst{Hist}(A)}^E_t \land E' = \sem{\mathit{Eqs} \cdot G}^E_t \imp \sem{P}^{E'}_t
\]
\end{definition}
\footnote
{KLS: At this point we would like to assert that the AGREE ``consistency
 check'' for a contract implies code contract correctness.}

The evaluation semantics given above is based on the stream
computation model, where values `deep in the past' can be accessed in
computations. However, the conventional program languages we wish to
target do not support such a view. Thus code generation is only
defined for a class of \emph{well formed} equations.

\begin{definition}[Wellformedness]

\end{definition}

\subsubsection{Example}
